---
title: "Favourite foods"
author: "Annie Somerville"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: yes
    toc_float: yes
---

```{r load-packages, message=FALSE}
library(tidyverse)
library(readxl)
```

## Exercise 1

* Read in the Excel file called `favourite-food.xlsx` from the `data-raw/` folder.

```{r fav-food}
fav_food <- read_excel("data-raw/favourite-food.xlsx")
fav_food
```

* Clean up `NA`s and make sure you're happy with variable types.

```{r clean-up}
fav_food_cleaned_names <- fav_food %>%
  janitor::clean_names()
names(fav_food_cleaned_names)

write.csv(fav_food_cleaned_names, file = 'data-raw/fav-food-cleaned-names.csv')

fav_food_cleaned_names <- read_csv(file = "data-raw/fav-food-cleaned-names.csv",
         na = c("N/A", "99999"))
```

``` {r five}
fav_food_cleaned_names <- fav_food_cleaned_names %>%
  mutate(
    age = case_when(
      full_name == "Chidiegwu Dunkel" ~ "5",
      TRUE                            ~ age
       ),
    age = as.numeric(age)
    )

fav_food_cleaned_names
```

* Convert SES (socio economic status) to a factor variables with levels in the 
following order: `Low`, `Middle`, `High`.

```{r factor}
fav_food_cleaned_names$ses <- as.factor(fav_food_cleaned_names$ses)

class(fav_food_cleaned_names$ses)
```

* Write out the resulting data frame to `favourite-food.csv` in the `data/` folder.

```{r write-out}
write.csv(fav_food_cleaned_names, file = 'data-raw/favourite-food.csv')
```

* Finally, read `favourite-food.csv` back in from the `data/` folder and 
observe the variable types. Are they as you left them?

```{r read-in}
fav_food_csv <- read.csv("data-raw/favourite-food.csv")
fav_food_csv
```

## Exercise 2

* Read in the Excel file called `favourite-food.xlsx` from the `data-raw/` folder, 
taking care of `NA`s and variable types.

```{r read-in-2}
fav_food2 <- read_excel("data-raw/favourite-food.xlsx")

fav_food_cleaned_names2 <- fav_food2 %>%
  janitor::clean_names()
names(fav_food_cleaned_names2)

write.csv(fav_food_cleaned_names2, file = 'data-raw/fav-food-cleaned-names2.csv')

fav_food_cleaned_names2 <- read_csv(file = "data-raw/fav-food-cleaned-names2.csv",
         na = c("N/A", "99999"))

fav_food_cleaned_names2 <- fav_food_cleaned_names2 %>%
  mutate(
    age = case_when(
      full_name == "Chidiegwu Dunkel" ~ "5",
      TRUE                            ~ age
       ),
    age = as.numeric(age)
    )

fav_food_cleaned_names2
```

* Convert SES (socio economic status) to a factor variables with levels in the 
following order: `Low`, `Middle`, `High`.

```{r factor2}
fav_food_cleaned_names2$ses <- as.factor(fav_food_cleaned_names2$ses)

class(fav_food_cleaned_names2$ses)
```

* Write out the resulting data frame to `favourite-food.rds` in the `data/` folder.

```{r write-rds}
write_rds(fav_food_cleaned_names2, file = 'data-raw/favourite-foods.rds')
```

* Read `favourite-food.rds` back in from the `data/` folder and observe the 
variable types. Are they as you left them?

```{r read-rds}
fav_food_rds <- read_rds("data-raw/favourite-foods.rds")
fav_food_rds
```

## References {#references}
1. Assignment Adapted from [Mine Cetinkaya-Rundel's Data Science in a Box](https://github.com/rstudio-education/datascience-box)